Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-17 appendix.qmd
Updated PBMC3k data
Updated PBMC3k data figure link

As we were writing this paper @xia2023 appeared proposing a new method called scDEED helping to assess the validity of a \gD{} embedding. scDEED calculates a reliability score for each cell embedding based on the similarity between the cell’s \gD{} embedding neighbors and its neighbors prior to embedding. A low reliability score suggests a dubious embedding. It can help in the deciding on optimal hyper-parameters. Here we illustrate how our method compares with the results from scDEED

@@

@fig-model-pbmc-author-proj-scdeed shows layout from the original published article along with that suggested by scDEED to be more accurate. The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.

>>>>

As we were writing this paper @xia2023 appeared proposing a new method called scDEED helping to assess the validity of a \gD{} embedding. scDEED calculates a reliability score for each cell embedding based on the similarity between the cell’s \gD{} embedding neighbors and its neighbors prior to embedding. A low reliability score suggests a dubious embedding. It can help in the deciding on optimal hyper-parameters. Here we illustrate how our method compares with the results from scDEED.

@@

@fig-pbmc-mse-umap shows layout from the original published article along with that suggested by scDEED to be more accurate. The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.
